Sexual selection, environmental variation, range size and speciation in passerine

### 2018-10-02 Manuscript.Rmd
Critical Analysis and index add

---
title: "Sexual selection, environmental variation, range size and speciation in passerine birds"
author: 
- Justin G. Cally*\textsuperscript{§}, Iliana Medina Guzmán\textsuperscript{§}, Luke Holman\textsuperscript{§} and Devi Stuart-Fox\textsuperscript{§}
- 
- '*\textit{justin.g.cally@gmail.com}'
output:
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
  pdf_document: default
bibliography: References/SS_Extinction.bib
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual Selection and Extinction} \usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\doublespacing \usepackage[font={small}]{caption}
csl: References/AmNat.csl
abstract: Sexual selection may influence patterns of diversification. Yet their relationship remains unclear with controversy on whether sexual selection leads to a net increase or decrease in speciation rate. Sexual selection is predicted to increase speciation rate as it can facilitate adaptation to new environments by elevating the average fitness of a population. Additionally, sexual selection may work alongide divergent natural selection and swiflty promote reproductive isolation. Here we use sexual dichromatism as a proxy for sexual selection across 5,812 species of passerine birds and a multivariate male-bias measure of sexual selection (sexual dimorphism, polygyny and paternal care) across 2,465 species to investigate the association of sexual selection with speciation and extinction. Importantly, we test whether the effects of sexual selection on speciation are dependent upon ecological pressures such as temperature variability across time and space, primary productivity and range size. Our results from three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) find that male-bias sexual selection but not sexual dichromatism predicts speciation. Additionally, we find that birds with smaller ranges have higher speciation rates; however the causality of this finding remains inconclusive.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

\newpage
#Introduction xxxxxx

**Process of speciation**: Selection pressures drive speciation. As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011]. In this view populations will speciate and adapt more readily to to new divergent environments if the selection pressures and evolvability of the traits under selection are high. Hence, the rate at which speciation occurs is likely dependent on both the populations adaptive potential/fitness and the evolvability **mmm here it becomes unclear because you are suggesting the link between speciation and the evolvability of traits, but is not clear what you mean by evolvability: a. the potential to evolve (e.g. genetic architecture) b. the rate of trait evolution or c. the magnitud of the difference in a trait between populations. I would either define evolvability earlier or replace the word with something easier to grasp** of phenotypic traits such as morphology or chromatism **color? avoid jargon when possible** [@Rabosky_morph_2013; @Hugall_2012]. Sexual selection is an evolutionary force likely to affect fitness components as well as the evolvability **here I would just use 'evolution' rather than evolvability** of morphological and chromatic traits such as ornaments involved in mating displays or weapons used in male-male competition. It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, speciation and extinction.

**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and **the strength of sexual selection**can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013]. Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations, [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations in a population [@Whitlock_2000]. Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation. However, **sexual selection can increase the fitness of a population when ** the genotypes under sexual selection (e.g. those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (e.g. female fecundity or survival). This positive correlation is unlikely to always be the case; regularly, sexual selection favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006]. For instance, in species with high levels of sexual selection mating success becomes more variable and uncertain, thus individuals must heavily invest to increase their chances of successfully reproducing. Through prioritising mating success a sex may invest in costly weaponry and sexual signals, while forgoing strategies that would improve offspring fitness (parental care). This is a potentially costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].

**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate ecological diversification ***not sure what you mean by this, like specialization in different niches? or generalization?** and speciation through increased evolvability of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982]. ** I think a example illustrating the above would be good** Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013]. But sexual selection is unlikely to be occurring without the effects of natural selection. In this situation, adaptation and speciation may be further enforced if female preferences in mating signals are dually **which are the two?** favoured by natural selection [@Maan_2008]. The extent to which sexually selected traits can evolve is dependent on the underlying genetic variation. Under sexual selection a wider phenotypic space may be available for a population to evolve into than natural selection acting alone, as genetic variation can be maintained under sexual conflict arising from heightened sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016]. Furthermore, the evolvability and transmission of fitness and sexually selected traits may vary due to different sex determination systems across taxa. In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are anticipated to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage coloration [@Hastings_1994; @Reeve_2003]. We might therefore expect birds with higher opportunities for sexual selection to have increased evolvability and speciation rates, especially in environments where natural selection pressures are high and working concordantly to sexual selection. *** I think that sentence needs clarification and a reference, by higher opportunities you mean stronger sex sel? ***

**Environmental interaction**: Importantly, the effects of sexual selection **on?** adaptation and speciation may depend on environmental conditions. Studies suggest that under stressful/changing environments sexual selection may have greater fitness benefits than under benign environments. In stressful environments, the positive correlation between male mating success and female fitness components improves, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017]. Conversely, under benign environments environmental stability ensures consistent selection, preferentially removing genetic variation at sexually concordant loci (i.e. loci where the optimum genotype is equivelant for both sexes). In these cases genetic variation preferentially remains at sexually antagonistic loci; causing displacement of male and female traits from their optimum. These predictions would ensue that populations **with stronger sexual selection would have increased fitness and adaptive benefits ** , allowing them to adapt to novel environments [*reviewed in* @Candolin_2008]. Increased environmental variability in a species ecological and climatic niche may also work synergestically with natural selection to promote diverging phenotypic traits **maybe leave out this last sentence? we are not really interested in how nat and sex sel interact to generate phenotypic variation but how combined they can affect speciation** [@Safran_2013]. In certain circumstances sexual selection will reinforce diverging locally adaptated phenotypes through mate choice on phenotypes that advertise quality in a specific ecological context [@Maan_2011]. **any examples?** If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981]. **maybe mention magic traits? those that are the result of natural selection but end up affecting sexual selection and thus speciation?** Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in sticklebacks [@Wong_2007; @Candolin_2007]. In this opposing view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.

**Previous limitations**: The association between sexual selection and speciation has been investigated using many macroevolutionary approaches across various taxa. Importantly, macroevolutionary comparative studies can examine the association between sexual selection and extinction risk in more diverse taxon, over longer evolutionary time-scales and in more natural environments. A meta-analysis of 64 effect sizes from 20 papers found a small, significant **positive?** association between sexual selecetion and speciation [@Kraaijeveld_2011]. However, variation in sexual selection proxies used and other methodological choices (species richness vs rate estimates) led to heterogeneity in effect sizes for the correlation between sexual selection and speciation. Given that species richness and speciation rate are not neccesarily correlated e.g. [@Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies. Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa, which arguably decreases the power to detect an association, especially if phylogenetic signal is strong. Secondlly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014]. Thirdly, until recently studies have rarely incorperated extensive bioclimatic predictors into large-scale comparative analyses [*sensu* @Jetz_env_2011; @Pigot_2016]. ** I would give more importance to this one and frame it slightly different: e.g. Until recently studies on the association between sexual selection and speciation have largely ignored the contribution of environmental variables to explain variation in speciation rates. Furthermore, these studies haven't taken into account either the potential interaction between natural and sexual selection in the generation of new species.** Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague the ability to draw widely accepted findings **I don't know if this is true**.

**Our study**: We chose to do our study on passerine birds because of the wealth of data available for these (Order: Passeriformes). Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset (n = 918) [@Armenta_2008]. Here we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$;) in 5,812 species of passerines. Moreover, we assess the effects of sexual dichromatism measures based on RGB values alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465).**the bit of the sex sel index looks a bit lost in this paragraph but it's quite important, so I'd move it up and bring more attention to it** Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).

@@

We examined the effect of sexual selection on speciation and extinction rate in 97 % of passerines (*n* = 5,812 species; 58 % of all birds). Specifically, we (i) compiled datasets for sexual dichromatism/selection strength and environmental variability, (ii) obtained estimates of speciation and extinction rates across passerines, and (iii) conducted phylogenetic generalized least-squares (PGLS) regressions. Many of the processes undertaken in this study are documented with reproducible code in the electronic supplementary material (ESM).

@@

We used sexual dichromatism as proxy of sexual selection strength. Male and female plumage scores have previously been measured across 5,983 species of passerines [@Dale_2015]. Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011]. The mean values of red, green and blue were calculated and provided male and female plumage scores. Here we use the absolute difference between males and female RGB scores as an estimate of sexual dichromatism. Additionally, we compiled a colour discriminability measure of sexual dichromatism from spectrophotometer measures [@Armenta_2008]. These measurements include varition in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxanomic samples (as opposed to colour drawings). However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis. While there is a correlation between these two measures, there is residual variation ($r = 0.79$; **Figure S9**).

@@

###Environmental variables 
Using expert range maps [@birdlife_2017] we obtained estimates of species range size. Using these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to. Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015] **was is literally manually? OMG! or do you have a code that does it really well? Can I have it!? :)** . From 1,000 random point samples in each range we extracted 19 present day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017]. From these values we obtained means and standard deviations for each species. Using the same spatial sampling we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006]. Furthermore, to estimate variability in the energy available to species we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution. Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005]. Using this data, which we provide as a potentially useful data resource (see ESM), we generated five predictors of speciation associated with different patterns in environmental variability (see below).

@@

Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014]. This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary regimes with variable speciation and extinction rates. Importantly, these models provide tip-rate estimates of speciation and extinction rate. The parameters of the 100 BAMM runs are detailed in full in the ESM; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014]. Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's. We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (Table S3, Table S4). Further details of BAMM parameters and output are available through ESM, with tip-rate means and variances included in the ESM. Additionally, given the variability in BAMM estimates we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on an MCC from a genetic-only phylogeny across all birds [@Harvey_2017]. All analyses were conducted on log-rates.

@@

Using the extracted environmental variables from each species range size we developed biologically meaningful predictors of environmental variation/stress relating to (i) seasonal climate variation, (ii) spatial climate variation and (iii) long-term climate variation. To obtain seasonal climate variation we used mean values of temperature seasonality (BIO4) for each range. To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature (BIO4) and precipitation seasonality (BIO15). PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (**Table S1**). Given that species range is a potentially informative predictor of speciation and extinction we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; **Figure S1**). To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values. Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates  (**Table S2**, **Figure S2**). Despite similar theoretical underpinnings the predictors of environmental variability used in PGLS analyses are not strongly correlated (**Figure S3**). Details and code to generate these predictors can be found within the ESM.

@@

#path analysis
best_model <- readRDS('data/path_model.rds')
```

##Effects of sexual dichromatism/selection on speciation are not dependent upon environmental variability

>>>>

---
title: Sexual selection, environmental variation, range size and speciation in passerine
  birds
author:
- Justin G. Cally*\textsuperscript{§}, Devi Stuart-Fox\textsuperscript{§}, Luke Holman\textsuperscript{§} and Iliana Medina Guzmán\textsuperscript{§}
- ~
- '*\textit{justin.g.cally@gmail.com}'
bibliography: References/SS_Extinction.bib
output:
  pdf_document: default
  html_document: default
  word_document:
    pandoc_args: --smart
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual Selection and Extinction} \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}
  \linenumbers \usepackage{setspace}\onehalfspacing \usepackage[font={small}]{caption}
  \usepackage{epigraph} \setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
  \setlength\beforeepigraphskip{1\baselineskip} \setlength\afterepigraphskip{2\baselineskip}
  \renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape}
  \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue}
csl: References/AmNat.csl
abstract: Sexual selection may influence patterns of diversification. Yet their relationship
  remains unclear with controversy on whether sexual selection leads to a net increase
  or decrease in speciation rate. Sexual selection is predicted to increase speciation
  rate as it can facilitate adaptation to new environments by elevating the average
  fitness of a population. Additionally, sexual selection may work alongide divergent
  natural selection and swiflty promote reproductive isolation. Here we use sexual
  dichromatism as a proxy for sexual selection across 5,812 species of passerine birds
  and a multivariate male-bias measure of sexual selection (sexual dimorphism, polygyny
  and paternal care) across 2,465 species to investigate the association of sexual
  selection with speciation and extinction. Importantly, we test whether the effects
  of sexual selection on speciation are dependent upon ecological pressures such as
  temperature variability across time and space, primary productivity and range size.
  Our results from three distinct measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$,
  $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) find that male-bias
  sexual selection but not sexual dichromatism predicts speciation. Additionally,
  we find that birds with smaller ranges have higher speciation rates; however the
  causality of this finding remains inconclusive.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

#Introduction

**Process of speciation**: Selection pressures drive speciation. As populations diverge they accumulate genetic and phenotypic changes that ultimately erect barriers against gene flow [@Presgraves_2010; *but see* @Shaw_2011]. In this view, populations will speciate and adapt more readily to to new divergent environments if selection pressures are high. Hence, the rate at which speciation occurs is likely dependent on both the populations response to selection (adaptation) and the potential for divergent phenotypic traits to evolve [*e.g.* morphology or colour, @Hugall_2012; @Rabosky_morph_2013]. Sexual selection is an evolutionary force likely to affect fitness components as well as the evolution of ornaments involved in mating displays or weapons used in male-male competition. It is therefore hypothesised that sexual selection is a key driver of diversification through shifting the rates of adaptation, divergence, speciation and extinction.

**Sexual selection and fitness**: Sexual selection is applied by one sex on the other - more often by females on males [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008] and the strength of sexual selection can increase or decrease the average fitness of a population; often termed population fitness [*reviewed in*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013]. Specifically, in line with the 'good genes' perspective, sexual selection may purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fixate beneficial mutations in a population [@Whitlock_2000]. Thus, sexual selection is thought to be a key process in mitigating extinction risk and facilitating adaptation. However, sexual selection can increase the fitness of a population when the genotypes under sexual selection (*e.g.* those that increase male mating/fertilisation success) are genetically positively correlated with traits affecting fitness (*e.g.* female fecundity or survival). This positive correlation is unlikely to always be the case; regularly, sexual selection favours traits having negative pleiotropic effects on female fitness [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006]. For instance, in species with high levels of sexual selection mating success becomes more variable and uncertain, thus individuals must heavily invest to increase their chances of successfully reproducing. Through prioritising mating success a sex may invest in costly weaponry and sexual signals, while forgoing strategies that would improve offspring fitness (parental care). This is a costly consequence of sexual selection that affects persistence, adaptation and extinction [@Kokko_Brooks_2003].

**Sexual selection, heritability and mechanisms of speciation**: Alongside changes to fitness, sexual selection has been proposed to facilitate speciation through the evolution of divergent signals/ornaments associated with improved mating success [@Lande_1981; @Lande_1982]. For instance, in antbirds (Thamnophilidae), genera with complex songs and stark dichromatism were more species rich [@Seddon_2008]. Under this model of speciation, divergent sexual signals in one sex (usually males) co-evolve with divergent preferences for those signals in the opposite sex, and by itself sexual selection could lead to behavioural reproductive isolation [@Safran_2013]. Concurrently, sexual selection can promote the extent to which populations diverge by enlarging the phenotypic space diverging species can evolve into. This prediction emanates from increased genetic diversity as a result of sexual conflict: a consequence of sexual selection [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016]. However, the genetic architecture (*per se* heritability) of fitness and sexually selected traits may vary due to different sex determination systems across taxa. In male homogametic species (ZZ/ZW or ZZ/ZO) --- such as birds --- males are anticipated to have increased heritability of male-specific genotypes on the Z chromosome associated with mating success: such as elaborate ornamentation or plumage colouration [@Hastings_1994; @Reeve_2003]. We might therefore expect birds with magnified sexual selection to have increased adaptation and speciation rates.

**Environmental interaction**: Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions. Studies suggest that under stressful/changing environments sexual selection may have greater fitness benefits than under benign environments. In stressful environments, the positive correlation between male mating success and female fitness components improves, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017]. Conversely, under benign environments environmental stability ensures consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivelant for both sexes). In these cases genetic variation preferentially remains at sexually antagonistic loci; causing displacement of male and female traits from their optimum. These predictions would ensue that populations with stronger sexual selection would have increased fitness and adaptive benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008]. In certain circumstances sexual selection will reinforce diverging locally adaptated phenotypes through mate choice on phenotypes that advertise quality in a specific ecological context [@Maan_2011]. If mate choice operates on both sexes --- or there is a correlation in trait values between a pair --- assortative mating is likely to strengthen this positive feedback loop and promote population divergence [@Lande_1981]. In several birds, beak morphology is a trait that arose under natural selection and is now a trait under sexual selection (a.k.a. "magic traits"). In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007]. Sexual selection may also interact with natural selection in opposing directions when environmental conditions modify the sexually selected signal, such as turbid waters that reduce the signalling strength in sticklebacks [@Wong_2007; @Candolin_2007]. In this opposing view, changing environments may limit the divergence of sexually selected traits and consequently reduce speciation rate.

**Previous limitations**: The association between sexual selection and speciation has been investigated using many macroevolutionary approaches across various taxa. Importantly, macroevolutionary comparative studies can examine the association between sexual selection and speciation in more diverse taxon, over longer evolutionary time-scales and in more natural environments. A meta-analysis of 64 effect sizes from 20 papers found a small, but positively significant association between sexual selecetion and speciation [@Kraaijeveld_2011]. However, variation in sexual selection proxies used and other methodological choices (species richness vs rate estimates) led to heterogeneity in effect sizes for the correlation between sexual selection and speciation. Given that species richness and speciation rate are not neccesarily correlated [*e.g.* @Rabosky_fish_2018] and the recent development and refinement of methodologies in obtaining tip-rate estimates of speciation and extinction [@Jetz_2012; @Rabosky_2014; @Rabosky_diversification_2018] the association between sexual selection and speciation rate is worth revisiting with methods that address several issues that have previously limited macroevolutionary comparative studies. Firstly, sexual selection strength in macroevolutionary studies are often estimated using measures of dimorphism or dichromatism; but these are often partial measures of sexual selection and may not account for the true variation in sexual selection across taxa; which arguably decreases the power to detect an association, especially if phylogenetic signal is strong. Secondlly, many studies use incomplete taxon sampling, with family-level analysis restricted to well sampled avian clades [@Seddon_2008; @Huang_2014]. Thirdly, until recently studies on the association between sexual selection and speciation have largely ignored the contribution of environmental variables to explain variation in speciation rates. Furthermore, these studies haven't taken into account either the potential interaction between natural and sexual selection in the generation of new species. Fourthly, phylogenetic uncertainty in the avian tree [@Jetz_2012; *but see* @Rubolini_2015] and uncertainty in tip-rate measures of speciation such as the diversification rate statistic ($\lambda_{DR}$) or node density ($\lambda_{ND}$) [@Rabosky_diversification_2018] as well as speciation and extinction tip-rates from models using BAMM (Bayesian Analysis of Macroevolutionary Mixtures) [*see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017] continue to plague our ability to draw widely accepted findings.

**Our study**: We chose to do our study on passerine birds because of the wealth of data available for these (Order: Passeriformes). Recently, @Huang_2014 found no association between sexual dichromatism and speciation using tip-rate estimates from a genetic-only phylogeny [@Jetz_2012] and sexual dichromatism data from a limited spectrophotometry dataset (n = 918) [@Armenta_2008]. But, here we assess the effects of sexual dichromatism measures based on RGB values (*n* = 5,812) alongside smaller datasets of sexual dichromatism from spectrophotometry measures (*n* = 581) and male-bias sexual selection measures (*n* = 2,465). Specifically, we assess the relationship between sexual dichromatism, environmental variability and their interaction with three tip-rate measures of speciation and one measure of extinction ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$ and $\mu_{BAMM}$). Using expert range maps and bioclimatic variables we are able to investigate whether the effect of sexual selection on speciation is are dependent on range size, seasonal variation in temperature, spatial temperature variation (across a species' range), long-term variation in temperature between now and the last inter-glacial (LIG) as well net primary productivity (NPP).

@@

We examined the effect of sexual selection on speciation and extinction rate in 97 % of passerines (*n* = 5,812 species; 58 % of all birds). Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions. Many of the processes undertaken in this study are documented with reproducible code in the electronic supplementary material (ESM).

@@

We used sexual dichromatism as proxy of sexual selection strength. Male and female plumage scores have previously been measured across 5,983 species of passerines [@Dale_2015]. Briefly, @Dale_2015 obtained sex-specific RGB (red, green and blue) values across six body patches (nape, crown, forhead, throat, upper breast and lower breast) from _Handbook of the Birds of the World_ (volumes 8–16) [@delhoyo2011]. The mean RGB values were calculated and provided male and female plumage scores. Here we use the absolute difference between males and female RGB scores as an estimate of sexual dichromatism. Additionally, we compiled a colour discriminability measure of sexual dichromatism from spectrophotometer measures [@Armenta_2008]. These measurements include varition in the ultraviolet and bird-visible range and --- unlike the RGB measures --- are sourced from taxanomic samples (as opposed to colour drawings). However, this spectrophotometry data is not as extensive as the RGB measures; with only 581 passerine species available for this analysis. While there is a correlation between these two measures, there is residual variation ($r = 0.79$; **Figure S9**).

@@

###Environmental variables

Using expert range maps [@birdlife_2017] we obtained estimates of species range size. From these distributions we were able to obtain estimates for extensive climatic conditions that each species is exposed to. Because of taxanomic changes to 1,230 species in the _Birdlife_ database [@delhoyo_2016] we manually matched these taxa with the names used in the sexual dichromatism dataset [@Dale_2015]. From 1,000 random point samples in each range we extracted 19 present-day bioclimatic variables (representing a variety of biologically relevent annual trends in temperature and precipitation) with 30 seconds (~1 km^2^) spatial resultion [@Fick_2017]. From these values we obtained means and standard deviations for each species. Using the same spatial sampling we extracted means and standard deviations of bioclimatic variables from the paleoclimate during the last interglacial (LIG; 120,000 - 140,000 years ago) [@Otto-Bliesner_2006]. Furthermore, to estimate variability in the energy available to species we obtained the mean and standard deviation of net primary productivity (NPP) values between 2000 - 2015 across each species distribution. Estimates of NPP had 30 second resolution and were obtained through MODIS (Moderate Resolution Imaging Spectroradiometer) primary production products stage 3 (MOD17A3) [@Zhao_2005]. Using this data, which we provide as a potentially useful data resource (see ESM), we generated five predictors of speciation associated with different patterns in environmental variability (see below).

@@

Secondly, we used BAMM to model the dynamics of speciation and extinction across the 100 phylogenetic trees [@Rabosky_2014]. This software uses a Bayesian approach (reversible-jump Markov Chain Monte Carlo) to generate a probability distribution of evolutionary regimes with variable speciation and extinction rates. Importantly, these models provide tip-rate estimates of speciation and extinction rate. The parameters of the 100 BAMM runs are detailed in full in the ESM; briefly we used a time-variable model with 100 expected number of shifts and prior rates set from the initial speciation and extinction values using the ``BAMMtools`` R package [@bammtools_2014]. Each BAMM model (one MCC and 100 random draws of the posterior) was run for 100 million generations, and given the computationally intensive nature of BAMM, runs were conducted across multiple CPU's. We obtained effective sample sizes for log-likelihood and number of rate shifts over 200 for each run (**Table S3**, **Table S4**). Further details of BAMM parameters and output are available through ESM, with tip-rate means and variances included in the ESM. Additionally, given the variability in BAMM estimates we also provide some analysis of BAMM shift configurations and tip-rate estimates from our run on the MCC tree and within a BAMM run on an MCC from a genetic-only phylogeny across all birds [@Harvey_2017]. All analyses were conducted on log-rates.

@@

Using the extracted environmental variables from each species range size we developed biologically meaningful predictors of environmental variation/stress relating to (*i*) seasonal climate variation, (*ii*) spatial climate variation and (*iii*) long-term climate variation. To obtain seasonal climate variation we used mean values of temperature seasonality (BIO4) for each range. To estimate levels of spatial environmental variation a species may endure we used the first principle component (PC1) from a PCA on standard deviations from all bioclimatic variables, excluding temperature (BIO4) and precipitation seasonality (BIO15). PC1 was heavily loaded towards bioclimatic variables relating to temperature, thus PC1 reflects the variation in temperature across a species' range (**Table S1**). Given that species range is a potentially informative predictor of speciation and extinction we controlled for the correlation between spatial variation and range size --- where larger ranges have larger variation in PC1 --- by taking the residuals of a fitted general additive model (GAM; **Figure S1**). To obtain long-term variation in climates for each species range we take the first principal component of the absolute difference in the bioclimatic variables between the LIG and current values. Similarly to spatial variation, the long-term climate variation is primarily loaded to temperature differences between the LIG and current climates  (**Table S2**, **Figure S2**). Despite similar theoretical underpinnings the predictors of environmental variability used in PGLS analyses are not strongly correlated (**Figure S3**). Details and code to generate these predictors can be found within the ESM.

@@

#path analysis
best_model <- readRDS('data/path_model.rds')
```
#Results

##Effects of sexual dichromatism/selection on speciation are not dependent on environmental variability

PGLS models with interaction terms between the respective measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation and NPP) were included in model selection. However, in all model selection phases for different measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and different measures of sexual selection (RGB values, spectrophotometry and male-biased sexual selection) no interaction terms were significant or present in the top models ($\delta$ AICc > 4; **Table S5**, **Table S6**, **Table S10**, **Table S13**). Hence we find no evidence that the effect of sexual selection on speciation or extinction is dependent upon certain environmental conditions. Furthermore, we find no support that these individual bioclimatic factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation and NPP) act as drivers of speciation independently from sexual dichromatism/selection (\autoref{model_results}, **Figure S10**).
